# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

* Load **activity.csv** and convert dates.
```{r echo=TRUE}
df_activity <- read.csv("activity.csv")
df_activity$date <- as.Date(df_activity$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?

* Create vector of the total number of steps per day.
```{r echo=TRUE}
v_total_steps_per_day <- tapply(df_activity$steps, df_activity$date,sum)
```

* Plot histogram of total number of steps per day.
```{r echo=TRUE}
hist(v_total_steps_per_day,
     xlab="Total Steps per Day", 
     ylab="Frequency", 
     main="Histogram of Total Steps per day",col="blue")
```

**Mean total steps taken per day:**
```{recho=TRUE}
mean(v_total_steps_per_day,na.rm=TRUE)
```

**Median total steps taken per day:**
```{r echo=TRUE}
median(v_total_steps_per_day,na.rm=TRUE)
```

## What is the average daily activity pattern?

* Create a data frame with the mean of steps over all days by time interval.
```{r echo=TRUE}
df_mean_steps_by_interval <- tapply(df_activity$steps,
                                    df_activity$interval, 
                                    mean,
                                    na.rm=TRUE)
```

* Create plot of the 5-minute interval and the average number of steps taken, averaged across all days.
```{r echo=TRUE}
plot(row.names(df_mean_steps_by_interval),
     df_mean_steps_by_interval,type="l",
     ylab="Mean # of Steps Taken (all Days)", 
     xlab="Time Intervals (5-minute)", 
     main="Average # of Steps Taken at 5-minute Intervals",
     col="blue")
```

* Determine the time interval that contains maximum average number of steps across all days.
```{r echo=TRUE}
interval_num <- which.max(df_mean_steps_by_interval)
interval_max_steps <- names(interval_num)
interval_num
```

**`r  interval_num `th interval, which is at time value `r  interval_max_steps `** 



## Imputing missing values

* Compute the number of NA values in the activity dataset.
```{r echo=TRUE}
num_na_values <- sum(is.na(df_activity))
num_na_values #Print number of NA values
```

* Fill in missing values using the average interval value across all days.
```{r echo=TRUE}
v_na_indices <-  which(is.na(df_activity))
v_imputed_values <- df_mean_steps_by_interval[as.character(df_activity[v_na_indices,3])]
count = 0
names(v_imputed_values) <- v_na_indices
for (i in v_na_indices) {
    df_activity$steps[i] = v_imputed_values[as.character(i)]
    count = count + 1
}
 
cat("Total ",count, "NA values were replaced and ",sum(is.na(df_activity))," remain.\n\r") 
```


## Are there differences in activity patterns between weekdays and weekends?

* Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE}
v_days <- weekdays(df_activity$date)
df_activity$day_type <- ifelse(v_days == "Saturday" | v_days == "Sunday", 
                               "Weekend", 
                               "Weekday")
df_mean_steps_by_interval <- aggregate(df_activity$steps, 
                                       by=list(df_activity$interval, 
                                               df_activity$day_type),
                                       mean)
names(df_mean_steps_by_interval) <- c("interval","day_type","steps")
```

* Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r echo=TRUE}
library(lattice)
xyplot(steps~interval | day_type, 
       df_mean_steps_by_interval,
       type="l",
       layout=c(1,2),
       xlab="Interval",
       ylab = "Number of Steps")

```


